---
title: "R Notebook"
output: html_notebook
---


Given a cursory analysis of the data we see that both the race of the mother and the race of the father appears to be correlated with baby weight. We also know that smoking is correlated with race, it is very posible that race is a confounding variable.
In the data given to us values 1-5 specify different types of white so we will create a dummy variable where all values for race 1-5 are mapped to white, 6 is mapped to Mexican, 7 is mapped to black, and 8 is mapped to Asian. 9 and 99 refer to mothers of mixed race and unknown race respectively. Since mixed race does not differentiate between different racial mixes as a category it is not useful for this analysis so we remove both.
```{r}
data_race = data
F <- function(x) if (x<=5) "white" else if (x == 6) "mex" else if (x == 7) "black" else if (x == 8) "asian"
mother_race <- unlist(lapply(data$race, F))
```

```{r}
Remove9and99 = function(toRemove, from) from[toRemove < 9]

wt_mother_race = Remove9and99(data_race$race, data_race$wt)
```

To get a sense of the data we will show a box plot.
```{r}
boxplot(wt_mother_race~mother_race,xlab="mothers race",ylab="baby weight")
```
It is unclear if there is a meaningful difference in the means so we will run an ANOVA test. I chose to use ANOVA to test whether race has a statistically significant effect on baby weight because it can determine if several categories have statistically different means. ANOVA has 3 assumptions: independence of cases, normality, and homoscedasticity.

Independence of cases means that each measurment is independent of the other. We cannot test this but we can assume it by assuming that the data was well-collected.

Normality of the residuals means that the distribution of the residuals is normal.

First we calculate the residuals:
```{r}
asian_residuals = wt_mother_race[mother_race == 'asian'] - mean(wt_mother_race[mother_race == 'asian'])
hist(asian_residuals)
```
```{r}
white_residuals = wt_mother_race[mother_race == 'white'] - mean(wt_mother_race[mother_race == 'white'])
hist(white_residuals)
```
```{r}
qqnorm(white_residuals);qqline(white_residuals, col = 2)
```

```{r}
black_residuals = wt_mother_race[mother_race == 'black'] - mean(wt_mother_race[mother_race == 'black'])
hist(black_residuals)
```
```{r}
qqnorm(black_residuals);qqline(black_residuals, col = 2)
```

```{r}
mex_residuals = wt_mother_race[mother_race == 'mex'] - mean(wt_mother_race[mother_race == 'mex'])
hist(mex_residuals)
```
```{r}
qqnorm(mex_residuals);qqline(mex_residuals, col = 2)
```

Looking at the histograms and the qqnorm plots, all of the distributions appear Normal.

Homoscedasticity means that all the variances are the same, but this is clearly not true based upon the box plots. However, it is established that ANOVA is robust, provided that the largest variance is no more than 4 times the smallest variance (which is true).

```{r}
summary(aov(wt_mother_race~mother_race))
```
Because of the extreemly low p-value mother's race appears to be significant.

We do the same for the father's race:
```{r}
fathers_race <- unlist(lapply(data$drace, F))
```

```{r}
Remove9and99 = function(toRemove, from) from[toRemove < 9]

wt_fathers_race = Remove9and99(data_race$drace, data_race$wt)
```
To get a sense of the data we will show a box plot.
```{r}
boxplot(wt_fathers_race~fathers_race,xlab="fathers race",ylab="baby weight")
```
```{r}
summary(aov(wt_fathers_race~fathers_race))
```

As we would expect, the father's race is also relevent. We expect that the mechanism for this is genetic differences in the child. Ideally we would like to test the race of the child. We can do this by removing the babies of mixed parentage.
```{r}
getBabyRace = function(x,y) if ((x == y) && (x < 9) && (y < 9)) F(x)
baby_race <- unlist(mapply(getBabyRace, data$race, data$drace))

# remove unnecessary mother race info
wt_race = data$wt[data$race < 9]
drace = data$drace[data$race < 9]
mrace = data$race[data$race < 9]

# remove unnecessary father race info
wt_race = wt_race[drace < 9]
mrace = mrace[drace < 9]
drace = drace[drace < 9]

wt_race = wt_race[drace == mrace]
```
To get a sense of the data we will show a box plot.
```{r}
boxplot(wt_race~baby_race,xlab="baby race",ylab="baby weight")
```

```{r}
summary(aov(wt_race~baby_race))
```

The increase in p-value is due to the decrease in sample size as shown by taking a random sample of mother and father race data sets and finding their p-value. As shown by:
```{r}
summary(aov(sample(wt_fathers_race, length(wt_race))~sample(fathers_race, length(baby_race))))
```

```{r}
summary(aov(sample(wt_mother_race, length(wt_race))~sample(mother_race, length(baby_race))))
```

--------------------------------------------------------------------------------------------------------------------------
By multiple metrics, race seems to be correlated with baby weight. Now we will test whether race is correlated with smoking in our data.
In order to determine this we need a sensible metric for smoking. This data set breaks the smoking data into 3 columns: a binary variable that determines whether a person has smoked, a time variable that determines how long ago a person quit if they formerly smoked, and a number variable representing how many cigarettes the individual smokes per day. Ideally I would use use the number column for only current and never smokers to maintain the numerical status of the smoking data. However this data is not Normal.
```{r}
smokedOrNo = data$smoke < 2
```
We then construct race variables that include only never and current smokers.
```{r}
getBabyRace = function(x,y) if ((x == y) && (x < 9) && (y < 9)) F(x)
baby_race <- unlist(mapply(getBabyRace, data$race[smokedOrNo], data$drace[smokedOrNo]))

# remove uncertain smokers
smoke = data$number[smokedOrNo]
smokeBinary = data$smoke[smokedOrNo]
drace = data$drace[smokedOrNo]
mrace = data$race[smokedOrNo]

# remove unnecessary mother race info
smoke = smoke[mrace < 9]
smokeBinary = smokeBinary[mrace < 9]
drace = drace[mrace < 9]
mrace = mrace[mrace < 9]

# remove unnecessary father race info
smoke = smoke[drace < 9]
smokeBinary = smokeBinary[drace < 9]
mrace = mrace[drace < 9]
drace = drace[drace < 9]

smoke_race = smoke[drace == mrace]
smokeBinary = smokeBinary[drace == mrace]
```
We run a box plot to visualize said data:
```{r}
boxplot(smoke_race~baby_race,xlab="baby race",ylab="number cigs")
```

So instead I will use simply use the chi-squared test to compare race to smoking while pregnent or never smoking. The assumptions of chi-squared are independence and sample size. Independence is assumed as explaned above. Sample size must be larger than 30, which is true for all of them.


```{r}
tbl = table(smokeBinary,baby_race)
tbl
chisq.test(tbl) 
```

While less significant than the relationship between race and baby weight it is still very significant at p < 0.05. Therefore, it is posible for race to be a confounding variable.
